On the Physics of Radio Halos in Galaxy Clusters: Scaling Relations and Luminosity Functions

### 2013-10-05 paper2.tex
very last Fabio

\documentclass[useAMS,usenatbib]{mn2e}                                  
%
%\def\del#1{{\bf  [deleted text] }} \def\new#1{{\bf #1}} \def\remark#1{{\bf  [remark: #1]}}
%\def\del#1{} \def\new#1{#1} \def\remark#1{}

\def\del#1{{}}
%\def\del#1{{\bf (DELETED TEXT)}}
\def\C#1{{\bf #1}}
%\def\C#1{#1}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf The underlying physics of giant and mini radio halos in galaxy clusters
    is still an open question. We find that mini halos (such as in Perseus and
    Ophiuchus) can be explained by radio-emitting electrons that are generated
    in hadronic cosmic ray (CR) interactions with protons of the intra-cluster
    medium. By contrast, the hadronic model either fails to explain the extended
    emission of giant radio halos (as in Coma at low frequencies) or would
    require a flat CR profile, which can be realized through outward streaming
    and diffusion of CRs (in Coma and A2163 at 1.4~GHz).  We suggest that a
    second, leptonic component could be responsible for the missing flux in the
    outer parts of giant halos within a new hybrid scenario and we describe its
    possible observational consequences.  To study the hadronic emission
    component of the radio halo population statistically, we use a cosmological
    mock galaxy cluster catalog built from the MultiDark simulation. Because of
    the properties of CR streaming and the different scalings of the X-ray
    luminosity ($L_{\rmn{X}}$) and the Sunyaev-Zel'dovich flux ($Y$) with gas
    density, our model can simultaneously reproduce the observed bimodality of
    radio-loud and radio-quiet clusters at the same $L_{\rmn{X}}$ as well as the
    unimodal distribution of radio-halo luminosity versus $Y$; thereby
    suggesting a physical solution to this apparent contradiction. We predict
    radio halo emission down to the mass scale of galaxy groups, which
    highlights the unique prospects for low-frequency radio surveys (such as the
    LOFAR Tier 1 survey) to increase the number of detected radio halos by at
    least an order of magnitude.}
\end{abstract}

@@

{\bf The presence of large-scale diffuse radio synchrotron emission in clusters
  of galaxies proves the existence of relativistic electrons and magnetic fields
  permeating the intra-cluster medium (ICM). This diffuse cluster radio emission
  can be observationally classified into two phenomena: peripheral radio relics,
  which show irregular morphology and polarized emission and appear to trace
  merger and accretion shocks, as well as radio halos (see, e.g.,
  \citealp{2012A&ARv..20...54F}).  Radio \mbox{(mini-)}halos (RHs) are
  characterized by unpolarized radio emission, are centered on clusters and
  show a regular morphology, resembling the morphology of the thermal X-ray
  emission. However, the short cooling length of synchrotron-emitting electrons
  at GHz frequencies ($\lesssim 100$~kpc) challenges theoretical models to
  explain the large-scale radio emission that extends over several Mpcs and
  calls for an efficient in-site acceleration process of electrons.}

@@

\begin{table} 
\begin{center}
\caption{Radio-halo and mini-halo characteristics.}
\medskip
%\begin{tabular}{cccccccc}
%\hline
%\phantom{\Big|}
%name & $z$ & $D$ & $\Delta d$ & $L_{1.4~\rmn{GHz},~\rmn{obs}}$ & model parameters & $L_{1.4~\rmn{GHz},~\rmn{model}}$ & references \\
%\phantom{\Big|}
%           &   & [$h_{70}^{-1}$~Mpc] & [$h_{70}^{-1}$~Mpc] & $10^{31}$ [$h_{70}^{-2}$~erg~s$^{-1}$~Hz$^{-1}$] & $\gamma_{\rmn{tu}}$, $\alpha_{\rmn{B}}$ & $10^{31}$ [$h_{70}^{-2}$~erg~s$^{-1}$~Hz$^{-1}$] & \\
%\hline \\[-0.5em]
%Coma           & $0.023$ & $101$ & $2.15$ & $0.72$  & 1, 0.6  & 0.86 &  [1, 2, 3]   \\
%               &         &       &        &                        & 4, 0.3  & 0.90  &  \\
%A2163         & $0.203$ & $962$ & $2.07$ & $15.36$     & 1, 0.3  & 13.43  &  [3, 4]  \\
%\hline \\[-0.5em]
%Perseus        & $0.018$ & $78$   & $0.15$ & $4.40$ & 3, 0.4   & 4.80 &  [3, 5, 6]  \\
%               &         &        &        &                       & 100, 0.3 & 3.97 &  \\
%Ophiuchucs     & $0.028$ & $121$  & $0.41$ & $0.19$       & 5, 0.7   & 0.19  &  [3, 4] \\
%               &         &        &        &                       & 100, 0.3 & 0.23 &   \\[0.5em]
\begin{tabular}{lcrrc}
\hline
\phantom{\Big|}
 cluster & $z$ & $D$~~ & $L_{1.4~\rmn{GHz}}$ & references \\
\hline \\[-0.5em]
Coma           & $0.023$ & $101$ & $0.72$  &  [1]   \\
A2163         & $0.203$ & $962$ & $15.36$  &  [2]  \\
\hline \\[-0.5em]
Perseus        & $0.018$ & $78$   & $4.40$ &  [3]  \\
Ophiuchucs     & $0.028$ & $121$  & $0.19$  &  [2] \\[0.5em]
\hline
\end{tabular}
\label{tab:RadioHalos}
\end{center}
\footnotesize{Note. Top two rows correspond to giant radio halos, while the
  bottom two rows are radio mini-halos.  $D$ is the luminosity distance in units
  of $h_{70}^{-1}$~Mpc and $L_{1.4~\rmn{GHz}}$ is the observed radio luminosity
  at 1.4~GHz in units of $10^{31}$~$h_{70}^{-2}$~erg~s$^{-1}$~Hz$^{-1}$.
  References: [1] \cite{1997A&A...321...55D} [2] \cite{2009A&A...499..679M} [3]
  \cite{1990MNRAS.246..477P}.}
\end{table}

@@

\begin{figure*}
\centering
%\includegraphics[width=0.75\textwidth]{figures/SB_profiles_ALL.eps}
%
\includegraphics[width=5.5cm,height=5.5cm,keepaspectratio]{figures/SB_Coma.eps}
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/XCR_Coma.eps}
\includegraphics[width=5cm,height=4.6cm]{figures/ProbComa.eps}
%%
\includegraphics[width=5.5cm,height=5.5cm,keepaspectratio]{figures/SB_A2163.eps}
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/XCR_A2163.eps}
\includegraphics[width=5cm,height=4.6cm]{figures/ProbA2163.eps}
%%
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/SB_Perseus.eps}
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/XCR_Perseus.eps}
\includegraphics[width=5cm,height=4.6cm]{figures/ProbPerseus.eps}
%%
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/SB_Ophiuchus.eps}
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/XCR_Ophiuchus.eps}
\includegraphics[width=5cm,height=4.6cm]{figures/ProbOphiuchus.eps}
%
\caption{Surface brightness modeling of the RHs in Coma, Abell~2163, Perseus and
  Ophiuchus. The left and middle panels show the RHs' azimuthally averaged
  surface brightness profiles and the corresponding CR-to-thermal pressure
  profiles $X_{\rmn{CR}}(r)$, respectively. Representative hadronic model
  parameters that fit the data well (solid) are compared to parameter choices
  that will be used in the second part of the paper (dashed, Sects.~\ref{sec:4}
  and \ref{sec:5}), which addresses RH statistics. While radio mini-halos can be
  fit by either set of parameters, for the latter choice of parameters, the
  hadronic model is not able to explain the emission in the outer parts of giant
  radio halos and would need a secondary, leptonic component (see text for
  details). This is exemplified in the lower left panels for Coma and Abell~2163
  that show the fraction of missing surface brightness for these parameter
  choices. In the middle panels, we additionally mark the RHs' radial extension
  by a vertical line. The panels on the right show the reduced-$\chi^2$ values
  of our model fits to the data in the $\gamma_{\rmn{tu}}-\alpha_{B}$ parameter
  space.  Regions of parameter space with reduced-$\chi^2$ values substantially
  larger than unity are excluded by the data while values much smaller than
  unity may point to an overestimate of the uncertainty intervals. Note that
  different parameter values that yield almost the same surface brightness
  profiles may result in very different $X_{\rmn{CR}}$ profiles. In the case of
  Abell~2163 and Ophiuchus, we adopt a 10 per cent uncertainty range instead of
  the errors reported by \citet{2009A&A...499..679M} to account for additional
  systematic uncertainties, e.g., residual point source contamination. For
  Perseus, we show the mini-halo data only for the range that is unaffected by
  residual point sources \citep{1990MNRAS.246..477P} and adopt a 10 per cent
  uncertainty budget.}
\label{fig:SBmodeling}
\end{figure*}

>>>>

\documentclass[useAMS,usenatbib]{mn2e}

\def\del#1{{}}
\def\C#1{{\bf #1}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  {\bf The underlying physics of giant and mini radio halos in galaxy clusters
    is still an open question. We find that mini halos (such as in Perseus and
    Ophiuchus) can be explained by radio-emitting electrons that are generated
    in hadronic cosmic ray (CR) interactions with protons of the intracluster
    medium. By contrast, the hadronic model either fails to explain the extended
    emission of giant radio halos (as in Coma at low frequencies) or would
    require a flat CR profile, which can be realized through outward streaming
    and diffusion of CRs (in Coma and A2163 at 1.4~GHz).  We suggest that a
    second, leptonic component could be responsible for the missing flux in the
    outer parts of giant halos within a new hybrid scenario and we describe its
    possible observational consequences.  To study the hadronic emission
    component of the radio halo population statistically, we use a cosmological
    mock galaxy cluster catalog built from the MultiDark simulation. Because of
    the properties of CR streaming and the different scalings of the X-ray
    luminosity ($L_{\rmn{X}}$) and the Sunyaev-Zel'dovich flux ($Y$) with gas
    density, our model can simultaneously reproduce the observed bimodality of
    radio-loud and radio-quiet clusters at the same $L_{\rmn{X}}$ as well as the
    unimodal distribution of radio-halo luminosity versus $Y$; thereby
    suggesting a physical solution to this apparent contradiction. We predict
    radio halo emission down to the mass scale of galaxy groups, which
    highlights the unique prospects for low-frequency radio surveys (such as the
    LOFAR Tier 1 survey) to increase the number of detected radio halos by at
    least an order of magnitude.}
\end{abstract}

@@

{\bf The presence of large-scale diffuse radio synchrotron emission in clusters
  of galaxies proves the existence of relativistic electrons and magnetic fields
  permeating the intracluster medium (ICM). This diffuse cluster radio emission
  can be observationally classified into two phenomena: peripheral radio relics,
  which show irregular morphology and polarized emission and appear to trace
  merger and accretion shocks, as well as radio halos (see, e.g.,
  \citealp{2012A&ARv..20...54F}).  Radio \mbox{(mini-)}halos (RHs) are
  characterized by unpolarized radio emission, are centered on clusters and
  show a regular morphology, resembling the morphology of the thermal X-ray
  emission. However, the short cooling length of synchrotron-emitting electrons
  at GHz frequencies ($\lesssim 100$~kpc) challenges theoretical models to
  explain the large-scale radio emission that extends over several Mpcs and
  calls for an efficient in-site acceleration process of electrons.}

@@

\begin{table} 
\begin{center}
\caption{Radio-halo and mini-halo characteristics.}
\medskip
\begin{tabular}{lcrrc}
\hline
\phantom{\Big|}
 cluster & $z$ & $D$~~ & $L_{1.4~\rmn{GHz}}$ & references \\
\hline \\[-0.5em]
Coma           & $0.023$ & $101$ & $0.72$  &  [1]   \\
A2163         & $0.203$ & $962$ & $15.36$  &  [2]  \\
\hline \\[-0.5em]
Perseus        & $0.018$ & $78$   & $4.40$ &  [3]  \\
Ophiuchucs     & $0.028$ & $121$  & $0.19$  &  [2] \\[0.5em]
\hline
\end{tabular}
\label{tab:RadioHalos}
\end{center}
\footnotesize{Note. Top two rows correspond to giant radio halos, while the
  bottom two rows are radio mini-halos.  $D$ is the luminosity distance in units
  of $h_{70}^{-1}$~Mpc and $L_{1.4~\rmn{GHz}}$ is the observed radio luminosity
  at 1.4~GHz in units of $10^{31}$~$h_{70}^{-2}$~erg~s$^{-1}$~Hz$^{-1}$.
  References: [1] \cite{1997A&A...321...55D} [2] \cite{2009A&A...499..679M} [3]
  \cite{1990MNRAS.246..477P}.}
\end{table}

@@

\begin{figure*}
\centering
%
\includegraphics[width=5.5cm,height=5.5cm,keepaspectratio]{figures/SB_Coma.eps}
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/XCR_Coma.eps}
\includegraphics[width=5cm,height=4.6cm]{figures/ProbComa.eps}
%%
\includegraphics[width=5.5cm,height=5.5cm,keepaspectratio]{figures/SB_A2163.eps}
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/XCR_A2163.eps}
\includegraphics[width=5cm,height=4.6cm]{figures/ProbA2163.eps}
%%
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/SB_Perseus.eps}
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/XCR_Perseus.eps}
\includegraphics[width=5cm,height=4.6cm]{figures/ProbPerseus.eps}
%%
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/SB_Ophiuchus.eps}
\includegraphics[width=5.6cm,height=5.6cm,keepaspectratio]{figures/XCR_Ophiuchus.eps}
\includegraphics[width=5cm,height=4.6cm]{figures/ProbOphiuchus.eps}
%
\caption{Surface brightness modeling of the RHs in Coma, Abell~2163, Perseus and
  Ophiuchus. The left and middle panels show the RHs' azimuthally averaged
  surface brightness profiles and the corresponding CR-to-thermal pressure
  profiles $X_{\rmn{CR}}(r)$, respectively. Representative hadronic model
  parameters that fit the data well (solid) are compared to parameter choices
  that will be used in the second part of the paper (dashed, Sects.~\ref{sec:4}
  and \ref{sec:5}), which addresses RH statistics. While radio mini-halos can be
  fit by either set of parameters, for the latter choice of parameters, the
  hadronic model is not able to explain the emission in the outer parts of giant
  radio halos and would need a secondary, leptonic component (see text for
  details). This is exemplified in the lower left panels for Coma and Abell~2163
  that show the fraction of missing surface brightness for these parameter
  choices. In the middle panels, we additionally mark the RHs' radial extension
  by a vertical line. The panels on the right show the reduced-$\chi^2$ values
  of our model fits to the data in the $\gamma_{\rmn{tu}}-\alpha_{B}$ parameter
  space.  Regions of parameter space with reduced-$\chi^2$ values substantially
  larger than unity are excluded by the data while values much smaller than
  unity may point to an overestimate of the uncertainty intervals. Note that
  different parameter values that yield almost the same surface brightness
  profiles may result in very different $X_{\rmn{CR}}$ profiles. In the case of
  Abell~2163 and Ophiuchus, we adopt a 10 per cent uncertainty range instead of
  the errors reported by \citet{2009A&A...499..679M} to account for additional
  systematic uncertainties, e.g., residual point source contamination. For
  Perseus, we show the mini-halo data only for the range that is unaffected by
  residual point sources \citep{1990MNRAS.246..477P} and adopt a 10 per cent
  uncertainty budget.}
\label{fig:SBmodeling}
\end{figure*}
